Hydrodynamics of two vesicles interacting via a Van der Waals potential

### 2019-01-07 vesicle_code/docs/adhesion/adhesion.tex
put in a few todos that ashley pointed out

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Vesicles (self-enclosed phopholipid bilayer membranes) have been widely utilized
as simple model membranes for cell mimics in biomedical and material engineering \cite{}. 
%
The ever expanding applications of vesicles have encouraged extensive investigations on vesicle hydrodynamics
and electrohydrodynamics.
Theoretical investigations are conducted to study the small-deformation or spheroidal dynamics of a nearly spherical  3D (or nearly circular 2D) vesicle in a fluid flow or under an external forcing \cite{Barthes-BieselRallison1981_JFM,Misbah2006_PRL,Vlahovska2007_PRE,Finken2008_EPL,ZhangZahnTanLin2013_PoF,Nganguia2013_PRE}. 
Various numerical schemes have been designed and deployed to simulate the hydrodynamics of many vesicles \cite{BagchiJohoson2005_JBE,Biben2005_EJP,Veerapaneni2009_JCP,SeolHuKimLai2016_JCP}.  
Experiments are performed to study vesicle hydrodynamics under various flowing conditions or forcing configurations \cite{MaderVitkova2006_EurPhysJE, Abkarian2008_SoftMatt,DahlNarsimhanGouveia2016_SoftMatt}. In a vesicle suspension under 
a hydrodynamic flow or an external field (such as an electric field) 
the interactions between
vesicles play a crucial role in determining the collective dynamics of
vesicle suspension. For example, boundary integral simulations show that
vesicles may align to form chains along the external electric field,
enhancing the suspension viscosity in the direction of alignment.
In numerical simulations of vesicle suspension, it is often assumed that lubrication layer between vesicle membranes act to keep
membrane-membrane interaction (such as Van der Waals interaction) negligible \cite{BagchiJohoson2005_JBE,Veerapaneni2009_JCP}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral Equation Formulation}
Because the fluid equations are linear and elliptic, an integral
equation formulation is possible.  This has several advantages,
including that only the vesicle interfaces require discretization.  We
start by defining the Stokes single-layer potential
\begin{align}
  \SS[\ff](\xx) &= \frac{1}{4\pi\mu} \int_\gamma \left(
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff(\yy) ds_\yy, \quad \xx \in \RR^2.
  \label{eqn:SLP}
\end{align}
The traction jump of the velocity $\uu(\xx) = \SS[\ff](\xx)$ is the
density function $\ff$, so the fluid velocity is $\xx \in \RR^2$ is
\begin{align*}
  \uu(\xx) = \uu_{\infty}(\xx) + \SS[\xxi](\xx).
\end{align*}
The traction jump is given by three terms: the bending, tension, and
adhesion.  Therefore
\begin{align*}
  \xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA \xx.
\end{align*}
Applying the no-slip boundary condition on each vesicle, the governing
equations are
\begin{align*}
  &\dot{\xx} = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \\
  &\xx_s \cdot \uu_s = 0, \\
  &\xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA\xx.
\end{align*}
Defining the bending operator as $\BB[\xx](\ff) = -\kappa_b \ff_{ssss}$,
and the tension operator $\TT[\xx](\sigma) = (\sigma \xx_s)_s$, the
no-slip boundary condition can be written as
\begin{align*}
  \pderiv{\xx}{t} = \SS \BB \xx + \SS \TT \sigma + \SS \AA \xx.
\end{align*}
We apply the IMEX-Euler method 
\begin{align*}
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \SS^N \BB^N \xx^{N+1} + 
  \SS^N \TT^N \sigma^{N+1} + \SS^N \AA^N \xx^N,
\end{align*}
that treats both the bending and tension terms
semi-implicitly~\cite{qua-bir2014}. Note that the adhesive term is
treated explicitly.  Therefore, a single time step requires solving
\begin{align}
  \xx^{N+1} - \Delta t \SS^N \BB^N \xx^{N+1} - 
    \Delta t \SS^N \TT^N \sigma^{N+1} = \xx^N + 
    \Delta t \SS^N \AA^N \xx^N,
  \label{eqn:IMEXEuler}
\end{align}
along with the inextensibility constraint that is discretized as
\begin{align*}
  \xx_s^{N} \cdot \xx_{s}^{N+1} = 1.
\end{align*}
We discretize the vesicles at a set of collocation points, compute the
bending and tension terms with Fourier differentiation, and apply
Alpert quadrature to the logarithmically singular single-layer potential
$\SS$~\cite{alp1999}.  The source and target points of the adhesion
force never coincide since they are always on different vesicles, so the
adhesion force~\eqref{eqn:adhesionForce} is computed with the spectrally
accurate trapezoid rule~\cite{tre-wei2014}.  To control the error and
achieve higher-order accuracy in time, we apply a time adaptive spectral
deferred correction method~\cite{qua-bir2016}.  The SDC method is able
to achieve second-order accuracy by iteratively applying the first-order
method~\eqref{eqn:IMEXEuler}.

@@

\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image01.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image02.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image03.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image04.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image05.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image06.png}
  \caption{Two vesicles in an extensional flow.  The reduced area is
  $0.8$, the adhesion strength is $0.7$, and the adhesion length scale
is $0.4$, extensional rate is $\chi = 0.05$.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi9em2_ra080_image01.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi9em2_ra080_image02.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi9em2_ra080_image03.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi9em2_ra080_image04.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi9em2_ra080_image05.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi9em2_ra080_image06.png}
  \caption{Two vesicles in an extensional flow.  The reduced area is
  $0.8$, the adhesion strength is $0.7$, and the adhesion length scale
is $0.4$, extensional rate is $\chi = 0.09$.}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\todo[inline]{$\sigma$ is overdefined as the Hamacher constant, and the
tension, and possibly the stress}

\section{Introduction}
Vesicles (self-enclosed phopholipid bilayer membranes) have been widely utilized
as simple model membranes for cell mimics in biomedical and material engineering \cite{}. 
%
The ever expanding applications of vesicles have encouraged extensive investigations on vesicle hydrodynamics
and electrohydrodynamics.
Theoretical investigations are conducted to study the small-deformation or spheroidal dynamics of a nearly spherical  3D (or nearly circular 2D) vesicle in a fluid flow or under an external forcing \cite{Barthes-BieselRallison1981_JFM,Misbah2006_PRL,Vlahovska2007_PRE,Finken2008_EPL,ZhangZahnTanLin2013_PoF,Nganguia2013_PRE}. 
Various numerical schemes have been designed and deployed to simulate the hydrodynamics of many vesicles \cite{BagchiJohoson2005_JBE,Biben2005_EJP,Veerapaneni2009_JCP,SeolHuKimLai2016_JCP}.  
Experiments are performed to study vesicle hydrodynamics under various flowing conditions or forcing configurations \cite{MaderVitkova2006_EurPhysJE, Abkarian2008_SoftMatt,DahlNarsimhanGouveia2016_SoftMatt}. In a vesicle suspension under 
a hydrodynamic flow or an external field (such as an electric field) 
the interactions between
vesicles play a crucial role in determining the collective dynamics of
vesicle suspension. For example, boundary integral simulations show that
vesicles may align to form chains along the external electric field,
enhancing the suspension viscosity in the direction of alignment.
In numerical simulations of vesicle suspension, it is often assumed that lubrication layer between vesicle membranes act to keep
membrane-membrane interaction (such as Van der Waals interaction) negligible \cite{BagchiJohoson2005_JBE,Veerapaneni2009_JCP}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral Equation Formulation}
Because the fluid equations are linear and elliptic, an integral
equation formulation is possible.  This has several advantages,
including that only the vesicle interfaces require discretization.  We
start by defining the Stokes single-layer potential
\begin{align}
  \SS[\ff](\xx) &= \frac{1}{4\pi\mu} \int_\gamma \left(
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff(\yy) ds_\yy, \quad \xx \in \RR^2.
  \label{eqn:SLP}
\end{align}
\todo[inline]{Define $\rr$ and other variables in
equation~\eqref{eqn:SLP}}
The traction jump of the velocity $\uu(\xx) = \SS[\ff](\xx)$ is the
density function $\ff$, so the fluid velocity is $\xx \in \RR^2$ is
\begin{align*}
  \uu(\xx) = \uu_{\infty}(\xx) + \SS[\xxi](\xx).
\end{align*}
The traction jump is given by three terms: the bending, tension, and
adhesion.  Therefore
\begin{align*}
  \xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA \xx.
\end{align*}
Applying the no-slip boundary condition on each vesicle, the governing
equations are
\begin{align*}
  &\dot{\xx} = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \\
  &\xx_s \cdot \uu_s = 0, \\
  &\xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA\xx.
\end{align*}
Defining the bending operator as $\BB[\xx](\ff) = -\kappa_b \ff_{ssss}$,
and the tension operator $\TT[\xx](\sigma) = (\sigma \xx_s)_s$, the
no-slip boundary condition can be written as
\begin{align*}
  \pderiv{\xx}{t} = \SS \BB \xx + \SS \TT \sigma + \SS \AA \xx.
\end{align*}
We apply the IMEX-Euler method 
\begin{align*}
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \SS^N \BB^N \xx^{N+1} + 
  \SS^N \TT^N \sigma^{N+1} + \SS^N \AA^N \xx^N,
\end{align*}
that treats both the bending and tension terms
semi-implicitly~\cite{qua-bir2014}. Note that the adhesive term is
treated explicitly.  Therefore, a single time step requires solving
\begin{align}
  \xx^{N+1} - \Delta t \SS^N \BB^N \xx^{N+1} - 
    \Delta t \SS^N \TT^N \sigma^{N+1} = \xx^N + 
    \Delta t \SS^N \AA^N \xx^N,
  \label{eqn:IMEXEuler}
\end{align}
along with the inextensibility constraint that is discretized as
\begin{align*}
  \xx_s^{N} \cdot \xx_{s}^{N+1} = 1.
\end{align*}
We discretize the vesicles at a set of collocation points, compute the
bending and tension terms with Fourier differentiation, and apply
Alpert quadrature to the logarithmically singular single-layer potential
$\SS$~\cite{alp1999}.  The source and target points of the adhesion
force never coincide since they are always on different vesicles, so the
adhesion force~\eqref{eqn:adhesionForce} is computed with the spectrally
accurate trapezoid rule~\cite{tre-wei2014}.  To control the error and
achieve higher-order accuracy in time, we apply a time adaptive spectral
deferred correction method~\cite{qua-bir2016}.  The SDC method is able
to achieve second-order accuracy by iteratively applying the first-order
method~\eqref{eqn:IMEXEuler}.

@@

\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi3em2_ra080_image01.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi3em2_ra080_image02.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi3em2_ra080_image03.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi3em2_ra080_image04.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi3em2_ra080_image05.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi3em2_ra080_image06.png}
  \caption{Two vesicles in an extensional flow.  The reduced area is
  $0.8$, the adhesion strength is $0.7$, and the adhesion length scale
is $0.4$, extensional rate is $\chi = 0.03$.}
\end{figure}

\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image01.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image02.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image03.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image04.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi9em2_ra080_image05.png}
  \includegraphics[width = 0.16\textwidth]{figs/extensional_adR4em1adS7em1Chi5em2_ra080_image06.png}
  \caption{Two vesicles in an extensional flow.  The reduced area is
  $0.8$, the adhesion strength is $0.7$, and the adhesion length scale
is $0.4$, extensional rate is $\chi = 0.05$.}
\end{figure}
